---
title       : Pe'ah Garden stats
subtitle    : Results for 2015
author      : Scott Gaul
job         : Pe'ah stats person
framework   : minimal      # 
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
ext_widgets : {rCharts: ["libraries/nvd3"]} 
mode        : selfcontained # {standalone, draft}
---

## 2015 overall results

We had great results again in 2014! 

We are on track for the biggest harvest ever in 2015, even more than [2014 which was the previous highest yields ever](http://sgaul.github.io/peah2014/).    

While we are still harvesting, gardeners so far have distributed 2,500 pounds of vegetables or about 230,000 calories - roughly three months of food for an adult. This is on track with last year's totals, which were the previous highest ever.

```{r setup, message = F, echo = F, cache = F}
require(rCharts)
knitr::opts_chunk$set(comment = NA, results = 'asis', tidy = F, message = T)
```


```{r chart1, echo = FALSE, message = FALSE, warning = FALSE}
library(ggplot2)
library(reshape)
library(plyr)
library(scales)
library(extrafont)
windowsFonts(Times=windowsFont("TT Times New Roman"))

#Load peah data
peah <- read.csv("Peah crops stats.csv")
names <- read.csv("names.csv")

n1 <- nPlot(y = 'Calories', x = 'Year', 
            width = 500, height = 300,
            data = ddply(peah, .(Year), summarise, 
                    Pounds = sum(Pounds),
                    Calories = sum(Calories)),
            type = 'discreteBarChart')
n1$xAxis(axisLabel='Year')
n1$yAxis(tickFormat = "#!function(d) {return d3.format(',.0f')(d)}!#", axisLabel='Calories')
n1$chart(forceY = c(0,170000), margin = list(left = 80))
n1$print('chart1')
```

Here is the same result in pounds harvested. 

```{r chart2, echo = FALSE, message = FALSE, warning = FALSE}
n1 <- nPlot(y = 'Pounds', x = 'Year', 
            width = 500, height = 300,
            data = ddply(peah, .(Year), summarise, 
                    Pounds = sum(Pounds),
                    Calories = sum(Calories)),
            type = 'discreteBarChart')
n1$xAxis(axisLabel='Year')
n1$yAxis(tickFormat = "#!function(d) {return d3.format(',.0f')(d)}!#", axisLabel='Pounds')
n1$chart(forceY = c(0,1700), margin = list(left = 80))
n1$print('chart2')
```

## 2015 individual results

We distributed 27 different kinds of vegetables this year - more than most years - with tomatoes, squash and zucchini yielding the most (over 300 lbs of each). 

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height=5, dev = "svg"}
peah15bycrop <- ddply(peah15, .(Crop), summarise, Pounds = sum(Pounds), Calories = sum(Calories))
peah15bycrop$Crop <- reorder(peah15bycrop$Crop, peah15bycrop$Pounds)

library(ggthemes)
ggplot(data = peah15bycrop) + 
  geom_bar(aes(x = Crop, y = Pounds), 
           stat = "identity", width =0.5, alpha =0.4) +
  geom_hline(yintercept=c(200,400,600), col="white", lwd=1) +
  labs(x = NULL, y = NULL) + 
  coord_flip() + 
  scale_y_continuous(labels = comma) + 
  theme_tufte(base_size = 16, ticks = F) + 
  annotate("text", x = "turnips", y = 400,  family="serif",
           label = c("Total pounds harvested\nfor each crop"))
```

The results by calories are similar, with a couple of high-calorie crops like beets and kale having more of an impact: 

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height=5, dev = "svg"}
peah15bycrop$Crop <- reorder(peah15bycrop$Crop, peah15bycrop$Calories)

library(ggthemes)
ggplot(data = peah15bycrop) + 
  geom_bar(aes(x = Crop, y = Calories), 
           stat = "identity", width =0.5, alpha =0.4) +
  geom_hline(yintercept=c(10000,20000,30000,40000), col="white", lwd=1) +
  labs(x = NULL, y = NULL) + 
  coord_flip() + 
  scale_y_continuous(labels = comma) + 
  theme_tufte(base_size = 16, ticks = F) + 
  annotate("text", x = "turnips", y = 30000,  family="serif",
           label = c("Total calories harvested\nfor each crop"))
```

The results by plot are here - congratulations to !

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height=5, dev = "svg"}
peah15byplot <- ddply(peah15, .(Name), summarise, Pounds = sum(Pounds), Calories = sum(Calories))
peah15byplot$Name <- reorder(peah15byplot$Name, peah15byplot$Calories)

ggplot(data = peah15byplot) + 
  geom_bar(aes(x = Name, y = Calories), 
           stat = "identity", width =0.5, alpha =0.4) +
  geom_hline(yintercept=c(0,50,100,150,200,250,300), col="white", lwd=1) +
  labs(x = NULL, y = NULL) + 
  coord_flip() + 
  scale_y_continuous(labels = comma) + 
  theme_tufte(base_size = 16, ticks = F) + 
  annotate("text", x = "Jen", y = 20000,  family="serif",
           label = c("Total calories harvested\nfor each plot"))
```


Individual results for the year are listed in the table below. 

```{r, results='asis', echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8}
#Current year results by plot and by crop
peah15 <- subset(peah, Year == 2015)
peah15 <- merge(peah15, names, by = "Plot.number")

library(xtable)
table <- ddply(peah15, .(Name, Crop), summarise, 
               Pounds = round(sum(Pounds),0), 
               Calories = round(sum(Calories),0))
print(xtable(table, digits = 0), type = "html", include.rownames = F)
```

The full data for all years can be found [here](https://docs.google.com/spreadsheet/ccc?key=0AlYsW526rxsmdDhIVzM0VDYzRkdLOXlvcldfQkJtcnc&usp=sharing). 

## Comparison with prior years

We had good results for virtually everything we planted and very good yields for squash, cucumbers, zucchini and tomatoes. Results for most other crops were above average. 

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height=12, fig.width = 10,dev = "svg"}
yearly <- ddply(peah, 
                .(Year, Crop), 
                summarise, 
                Pounds.per.plot = sum(Pounds.per.plot) / mean(Number.of.plots))

yearly <- subset(yearly, Crop %in% unique(subset(ddply(yearly, .(Crop), summarise, count = length(unique(Year))), count > 5)$Crop))

ggplot(data = yearly, aes(x = Year, y = Pounds.per.plot)) + 
  geom_line() + 
  scale_y_continuous(labels = comma) + 
  geom_point(data = subset(yearly, Year == 2015), 
             aes(x = Year, y = Pounds.per.plot), colour = "red") + 
  labs(title = "Crop harvest trends by year", y = "Pounds per plot") +
  facet_wrap(~ Crop, ncol = 3, scales = "free_y") + 
  theme_minimal(base_size = 16, base_family = "Times")
```

Why else did things go so well this year? 

We had more harvests from the garden than any prior year. Gardeners have harvested almost 400 times to date. Last year, we harvested 266 times and no other prior year had more than X harvests. (This may be partly due to better reporting on harvests in the last few years.)

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height=5, dev = "svg"}
library(ggthemes)
ggplot(data = ddply(peah, .(Year), summarise, count = length(Crop))) + 
  geom_bar(aes(x = Year, y = count), stat = "identity", width =0.5, alpha =0.4) +
  geom_hline(yintercept=c(50,100,150,200,250,300,350), col="white", lwd=1) +
  labs(x = NULL, y = NULL) + 
  scale_y_continuous(labels = comma, breaks=c(50,100,150,200,250,300,350)) + 
  theme_tufte(base_size = 16, ticks = F) + 
  annotate("text", x = 2006, y = 300,  family="serif",
           label = c("Number of harvests\nper year\n2002 - 2015"))

```

During the year, the most intense periods were at the end of June and July - gardeners harvested more than 250 pounds during each of these weeks. 

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height=5, dev = "svg"}
#Add week field to data
peah$Date <- as.Date(peah$Date, format = "%m/%d/%Y")
peah$Week = as.integer(format(strptime(peah$Date,"%Y-%m-%d"),"%W"))

ggplot(data = ddply(subset(peah, Year == 2015),
                    .(Week), 
                    summarise, 
                    Pounds = sum(Pounds, na.rm = T)), 
       aes(x = Month, y = Pounds)) + 
  geom_bar(aes(x = Week, y = Pounds), stat = "identity", width =.5, alpha =.4) +
  geom_hline(yintercept=c(100,200,300), col="white", lwd=1) +
  labs(x = "Week (number)", y = "Pounds harvested") + 
  theme_tufte(base_size = 16, ticks = F) + 
  annotate("text", x = 27, y = 275,  family="serif",
           label = c("Week of June 29")) + 
  annotate("text", x = 32, y = 300,  family="serif",
           label = c("Week of July 27"))
```

We visited the garden most often on Sunday, and least often on Saturdays, Mondays and Tuesdays. (Saturday visits may be after sundown or data entry errors on my part.)

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height=5, dev = "svg"}
#Summarize by day of week
peah$WEEKDAY = weekdays(as.Date(peah$Date))
peah$WEEKDAY <- factor(peah$WEEKDAY, levels = c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"))

ggplot(data = ddply(subset(peah, Year == 2015),
                    .(WEEKDAY), 
                    summarise, 
                    visits = length(Crop)), 
       aes(x = Month, y = visits)) + 
  geom_bar(aes(x = WEEKDAY, y = visits), stat ="identity", width =.5, alpha =.4) +
  geom_hline(yintercept=c(0,25,50,75,100), col="white", lwd=1) +
  labs(x = NULL, y = "Number of visits") + 
  theme_tufte(base_size = 16, ticks = F) + 
  annotate("text", x = "Thursday", y = 90,  family="serif",
           label = c("Frequency of visits\n(Days of the week)"))
```

